---
title: "Exploratory Analysis"
output: html_document
---

## Project title

Give your project a concise, interesting title that summarizes the entirety of your project

### Authors

Names of your team members

### Date

Winter 2022

### Abstract

No more than three sentences that summarize your project. Focus on the very most important aspects. For example: (1) "Our main question is .... This question is important because .... To address the question, we will ...." (2) "We are concerned with ..., because .... To address this concern, we plan to ...." (3) "Consider that .... This is important because .... Accordingly, we plan to ...."

### Keywords

3-5 keywords that summarize your project.
(e.g., "Keywords: human physiology; bicycle exercise; Gen Z; times-series data")

### Introduction  

Briefly introduce your project.  Include 3-5 research questions. What motivates the questions? Why are they important? (at least 200 words)

### Related Work  

Describe your topic and related work in this space. You must include 3 citations to related work (URLs to similar work, high quality articles from the popular press, research papers, etc. ) Please use a standard citation style of your choice. (at least 200 words)

### The Dataset

Where did you find the data? Please include a link to the data source  
Who collected the data?  
How was the data collected or generated?  
Why was the data collected?  
How many observations (rows) are in your data?  
How many features (columns) are in the data?  
What, if any, ethical questions or questions of power do you need to consider when working with this data?  
What are possible limitations or problems with this data?   (at least 200 words)

### Implications

Assuming you answer your research questions, briefly describe the expected or possible implications for technologists, designers, and policymakers. (at least 150 words)

### Limitations & Challenges
What challenges or limitations might you need to address with your project idea more broadly? Briefly discuss. (at least 150 words)

### Summary Information

Write a summary paragraph of findings that includes the 5 values calculated from your summary information R script

### Table

Include a table of aggregate information

Describe why you included the table and what information it reveals

### Chart 1

Include a chart

Describe why you chose this chart and what information it reveals

Here's an example of how to run an R script inside an RMarkdown file:

``````{r, echo = FALSE, code = readLines("chart1_example.R")}
```

### Chart 2

Include a chart

Describe why you chose this chart and what information it reveals

### Chart 3
```{r, echo = FALSE, code = readLines("Sally's chart.R")}
source("Sally's chart.R")
```

The line chart above shows the total vaccinations or saying total number of COVID-19 doses administered in each year from 2021 to 2023 for California, New York State, Pennsylvania, Washington, and American Samoa.

I choose the line plot to visualize the total vaccination data for the five states - California, New York State, Pennsylvania, Washington, and American Samoa - because it effectively reveals the trend and changes in the number of vaccinations administered over the years. The line plot is clearly showing the difference on the number of total vaccinations each states administered and it's a great tool for comparing between the data. It also easily to see the trend of the vaccinations through the line plot, whether it is increasing in a higher rate or lower rate through years.

Based on the plot, we can see California has the most total vaccinations of COVID in these five states and it's increasing in a high rate between year to year. Comparatively, the other four states are increasing at a lower rate than California. In these five states, we can see total vaccinations in American Samoa from the chart is almost stay in the value of 0, which is due to the number of total vaccinations here has a comparatively low number than the other states(e.g. total vaccination in California in 2021:816301
, total vaccination in American Samoa in 2021: 2124). From the plot, all of the total vaccinations in five states are increasing at a higher rate between 2021 to 2022 than 2022 to 2023.

The reason to choose these five states:
-- California: State with the highest number of COVID-19 cases
-- American Samoa: State with the lowest COVID-19 cases
-- New York State: State with most death due to the COVID-19
-- Pennsylvania: State with the highest death ratio of COVID-19
-- Washington: State with the earliest COVID-19 case